---
title: "Eqs. in Fig. 4"
author: "Masatoshi Katabuchi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 12pt
format:
  html:
    theme: coderpro
    toc: true
    toc-depth: 2
    number-sections: true
    smooth-scroll: true
    standalone: true
    embed-resources: true
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold"
)
```

```{r}
library(targets)
library(tidyverse)
```

```{r}
vaf_sp <- tar_read(fit5_summary_segments_noxylem_traits_sp_simple_log_vaf) |>
  filter(str_detect(variable, "beta\\[")) |>
  pull("q50")

vaf_seg <- tar_read(fit5_summary_segments_noxylem_traits_simple_log_vaf) |>
  filter(str_detect(variable, "beta\\[")) |>
  pull("q50")

ks_sp <- tar_read(fit5_summary_segments_noxylem_traits_sp_simple_log_ks) |>
  filter(str_detect(variable, "beta\\[")) |>
  pull("q50")

ks_seg <- tar_read(fit5_summary_segments_noxylem_traits_simple_log_ks) |>
  filter(str_detect(variable, "beta\\[")) |>
  pull("q50")

vaf_sp_a_int   <- vaf_sp[1] |> exp() |> round(2) |> format(nsmall = 2)  # beta[1,1]
vaf_sp_b_int   <- vaf_sp[2] |> round(3) |> format(nsmall = 3)            # beta[2,1]
vaf_sp_a_slope <- vaf_sp[3] |> round(3) |> format(nsmall = 3)            # beta[1,2]
vaf_sp_b_slope <- vaf_sp[4] |> round(3) |> format(nsmall = 3)            # beta[2,2]

vaf_seg_a_int   <- vaf_seg[1] |> exp() |> round(2) |> format(nsmall = 2)  # beta[1,1]
vaf_seg_b_int   <- vaf_seg[2] |> round(3) |> format(nsmall = 3)            # beta[2,1]
vaf_seg_a_slope <- vaf_seg[3] |> round(3) |> format(nsmall = 3)            # beta[1,2]
vaf_seg_b_slope <- vaf_seg[4] |> round(3) |> format(nsmall = 3)            # beta[2,2]

ks_sp_a_int   <- ks_sp[1] |> exp() |> round(2) |> format(nsmall = 2)  # beta[1,1]
ks_sp_b_int   <- ks_sp[2] |> round(3) |> format(nsmall = 3)            # beta[2,1]
ks_sp_a_slope <- ks_sp[3] |> round(3) |> format(nsmall = 3)            # beta[1,2]
ks_sp_b_slope <- ks_sp[4] |> round(3) |> format(nsmall = 3)            # beta[2,2]

ks_seg_a_int   <- ks_seg[1] |> exp() |> round(2) |> format(nsmall = 2)  # beta[1,1]
ks_seg_b_int   <- ks_seg[2] |> round(3) |> format(nsmall = 3)            # beta[2,1]
ks_seg_a_slope <- ks_seg[3] |> round(3) |> format(nsmall = 3)            # beta[1,2]
ks_seg_b_slope <- ks_seg[4] |> round(3) |> format(nsmall = 3)            # beta[2,2]
```



# Segments

## VAF

a = `r vaf_seg_a_int` VAF^ `r vaf_seg_a_slope`

b = `r vaf_seg_b_int` + `r vaf_seg_b_slope` Ln VAF

## Ks

a = `r ks_seg_a_int` Ks^ `r ks_seg_a_slope`

b = `r ks_seg_b_int` + `r ks_seg_b_slope` Ln Ks


# Species

## VAF

a = `r vaf_sp_a_int` VAF^ `r vaf_sp_a_slope`

b = `r vaf_sp_b_int` + `r vaf_sp_b_slope` Ln VAF

## Ks

a = `r ks_sp_a_int` Ks^ `r ks_sp_a_slope`

b = `r ks_sp_b_int` + `r ks_sp_b_slope` Ln Ks

# Trait scaling (mean and SD used for standardization)

```{r}
d <- read_csv("data/fd_k_traits.csv") |>
  filter(is.na(removed_k)) |>
  mutate(log_vaf = log(vaf), log_ks = log(ks))

seg <- d |>
  group_by(sample_id) |>
  summarise(log_vaf = mean(log_vaf, na.rm = TRUE),
            log_ks  = mean(log_ks,  na.rm = TRUE))

sp <- d |>
  group_by(species) |>
  summarise(log_vaf = mean(log_vaf, na.rm = TRUE),
            log_ks  = mean(log_ks,  na.rm = TRUE))

trait_scale <- tibble(
  level  = c("segment", "segment", "species", "species"),
  trait  = c("log_vaf", "log_ks",  "log_vaf", "log_ks"),
  mean   = c(mean(seg$log_vaf, na.rm = TRUE), mean(seg$log_ks, na.rm = TRUE),
             mean(sp$log_vaf,  na.rm = TRUE), mean(sp$log_ks,  na.rm = TRUE)),
  sd     = c(sd(seg$log_vaf,   na.rm = TRUE), sd(seg$log_ks,   na.rm = TRUE),
             sd(sp$log_vaf,    na.rm = TRUE), sd(sp$log_ks,    na.rm = TRUE))
)

knitr::kable(trait_scale, digits = 4,
             caption = "Mean and SD of log-transformed traits used for z-scoring")
```

# Back-transformed equations

```{r}
# z = (log(trait) - mu) / sigma
# log(a) = beta[1,1] + beta[1,2]*z
#         => a = exp(beta[1,1] - beta[1,2]*mu/sigma) * trait^(beta[1,2]/sigma)
# b = beta[2,1] + beta[2,2]*z
#     => b = (beta[2,1] - beta[2,2]*mu/sigma) + (beta[2,2]/sigma)*ln(trait)

bt <- function(beta_vec, mu, sigma) {
  list(
    a_const = exp(beta_vec[1] - beta_vec[3] * mu / sigma) |> round(2) |> format(nsmall = 2),
    a_power = (beta_vec[3] / sigma)                       |> round(3) |> format(nsmall = 3),
    b_const = (beta_vec[2] - beta_vec[4] * mu / sigma)   |> round(3) |> format(nsmall = 3),
    b_slope = (beta_vec[4] / sigma)                       |> round(3) |> format(nsmall = 3)
  )
}

pull_scale <- function(level_val, trait_val, col) {
  trait_scale |> filter(level == level_val, trait == trait_val) |> pull({{col}})
}

bt_vaf_seg <- bt(vaf_seg,
                 pull_scale("segment", "log_vaf", mean),
                 pull_scale("segment", "log_vaf", sd))
bt_ks_seg  <- bt(ks_seg,
                 pull_scale("segment", "log_ks",  mean),
                 pull_scale("segment", "log_ks",  sd))
bt_vaf_sp  <- bt(vaf_sp,
                 pull_scale("species", "log_vaf", mean),
                 pull_scale("species", "log_vaf", sd))
bt_ks_sp   <- bt(ks_sp,
                 pull_scale("species", "log_ks",  mean),
                 pull_scale("species", "log_ks",  sd))
```

## Segments

### VAF

a = `r bt_vaf_seg$a_const` VAF^ `r bt_vaf_seg$a_power`

b = `r bt_vaf_seg$b_const` + `r bt_vaf_seg$b_slope` Ln VAF

### Ks

a = `r bt_ks_seg$a_const` Ks^ `r bt_ks_seg$a_power`

b = `r bt_ks_seg$b_const` + `r bt_ks_seg$b_slope` Ln Ks

## Species

### VAF

a = `r bt_vaf_sp$a_const` VAF^ `r bt_vaf_sp$a_power`

b = `r bt_vaf_sp$b_const` + `r bt_vaf_sp$b_slope` Ln VAF

### Ks

a = `r bt_ks_sp$a_const` Ks^ `r bt_ks_sp$a_power`

b = `r bt_ks_sp$b_const` + `r bt_ks_sp$b_slope` Ln Ks
