---
title: "ab raito"
author: "Masatoshi Katabuchi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
# fontsize: 11pt
crossref:
  fig-title: Fig.
  fig-prefix: Fig.
format:
  html:
    theme: spacelab #readable #sandstone #spacelab #flatly
    toc: true
    toc-depth: 2
    toc-title: Contents
    # embed-resources: true
    # self_contained: true
    smooth-scroll: true
    highlight-style: github
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  cache = FALSE,
  echo = FALSE,
  fig.align = "center",
  fig.show = "hold",
  root.dir = rprojroot::find_root('_targets.R')
)
```


```{r}
library(targets)
library(tidyverse)
```

```{r, eval=FALSE}
hoge <- tar_manifest()

hoge |>
  filter(str_detect(name, "log_ks")) |>
  filter(str_detect(name, "summary")) |> pull(name)

hoge |>
  filter(str_detect(name, "species_ab_table_combined"))

data <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fd_k_traits_csv))

tar_read(stan_data_noxylem_log_ks)$uj |> dim()
tar_read(stan_data_noxylem_log_ks)$uj |>　rownames()

tar_read(stan_data_xylem_log_ks)$uj |> dim()
tar_read(stan_data_xylem_log_ks)$uj |>　rownames()

tar_read(fit_summary_segments_noxylem_traits_log_ks)
tar_read(fit2_summary_segments_xylem_traits_log_ks)


tar_read(fit_summary_segments_noxylem_traits_log_ks) |>
  filter(str_detect(variable, "^alpha_a")) |>
  filter(str_detect(variable, "15"))

tar_read(fit_summary_segments_noxylem_traits_log_ks) |>
  filter(str_detect(variable, "^alpha_b")) |>
  filter(str_detect(variable, "15"))

tar_read(fit_draws_segments_xylem_0.08) |>
  janitor::clean_names() |>
  dplyr::select(starts_with("alpha")) |>
  dplyr::select(ends_with("15"))

tar_read(sap_all_clean_0.08)$uj |> rownames()

```

sp no. 15
Use fit and fit2.

Ok, we use global a (hyper species) rather than species specific a.

Ks

- eq for a, ks, seg (xylem, fit2)
- eq for a, ks, sp (no xylem, fit)
- eq for b, ks, seg (xylem, fit2)
- eq for b, ks, sp (no xylem, fit)

- eq for a, vaf, seg (xylem, fit2)
- eq for a, vaf, sp (no xylem, fit)
- eq for b, vaf, seg (xylem, fit2)
- eq for b, vaf, sp (no xylem, fit)


## ks, no xylem (sp)
```{r}
summary_ks_sp <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_summary_segments_noxylem_traits_log_ks))
draws_ks_sp <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_draws_segments_noxylem_traits_log_ks)) |>
    janitor::clean_names()

ks_sp_a_int <- summary_ks_sp |>
  filter(variable == "beta_a[1,1]") |>
  pull(q50) |> exp() |> round(1)
ks_sp_a_power <- summary_ks_sp |>
  filter(variable == "beta_a[2,1]") |>
  pull(q50) |> round(3)
ks_sp_b_int <- summary_ks_sp |>
  filter(variable == "beta_b[1,1]") |>
  pull(q50) |> exp() |> round(3)
ks_sp_b_power <- summary_ks_sp |>
  filter(variable == "beta_b[2,1]") |>
  pull(q50) |> round(3)
```

$$
a = `r ks_sp_a_int` \times ^{`r ks_sp_a_power`}
$$
$$
b = `r ks_sp_b_int` \times K_s^{`r ks_sp_b_power`}
$$


## ks, xylem
```{r}

summary_ks_xylem <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_summary_segments_xylem_traits_log_ks))
draws_ks_xylem <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_draws_segments_xylem_traits_log_ks)) |>
    janitor::clean_names()

ks_xylem_a_int <- summary_ks_xylem |>
  filter(variable == "gamma_a[1,1]") |>
  pull(q50) |> exp() |> round(1)
ks_xylem_a_power <- summary_ks_xylem |>
  filter(variable == "gamma_a[2,1]") |>
  pull(q50) |> round(3)
ks_xylem_b_int <- summary_ks_xylem |>
  filter(variable == "gamma_b[1,1]") |>
  pull(q50) |> exp() |> round(3)
ks_xylem_b_power <- summary_ks_xylem |>
  filter(variable == "gamma_b[2,1]") |>
  pull(q50) |> round(3)
```

$$
a = `r ks_xylem_a_int` \times K_s^{`r ks_xylem_a_power`}
$$
$$
b = `r ks_xylem_b_int` \times K_s^{`r ks_xylem_b_power`}
$$


## vaf, no xylem (sp)
```{r}
summary_vaf_sp <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_summary_segments_noxylem_traits_log_vaf))
draws_vaf_sp <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_draws_segments_noxylem_traits_log_vaf)) |>
    janitor::clean_names()

vaf_sp_a_int <- summary_vaf_sp |>
  filter(variable == "beta_a[1,1]") |>
  pull(q50) |> exp() |> round(1)
vaf_sp_a_power <- summary_vaf_sp |>
  filter(variable == "beta_a[2,1]") |>
  pull(q50) |> round(3)
vaf_sp_b_int <- summary_vaf_sp |>
  filter(variable == "beta_b[1,1]") |>
  pull(q50) |> exp() |> round(3)
vaf_sp_b_power <- summary_vaf_sp |>
  filter(variable == "beta_b[2,1]") |>
  pull(q50) |> round(3)
```

$$
a = `r vaf_sp_a_int` \times \text{VAF}^{`r vaf_sp_a_power`}
$$
$$
b = `r vaf_sp_b_int` \times \text{VAF}^{`r vaf_sp_b_power`}
$$


## vaf, xylem
```{r}

summary_vaf_xylem <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_summary_segments_xylem_traits_log_vaf))
draws_vaf_xylem <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_draws_segments_xylem_traits_log_vaf)) |>
    janitor::clean_names()

vaf_xylem_a_int <- summary_vaf_xylem |>
  filter(variable == "gamma_a[1,1]") |>
  pull(q50) |> exp() |> round(1)
vaf_xylem_a_power <- summary_vaf_xylem |>
  filter(variable == "gamma_a[2,1]") |>
  pull(q50) |> round(3)
vaf_xylem_b_int <- summary_vaf_xylem |>
  filter(variable == "gamma_b[1,1]") |>
  pull(q50) |> exp() |> round(3)
vaf_xylem_b_power <- summary_vaf_xylem |>
  filter(variable == "gamma_b[2,1]") |>
  pull(q50) |> round(3)
```

$$
a = `r vaf_xylem_a_int` \times \text{VAF}^{`r vaf_xylem_a_power`}
$$
$$
b = `r vaf_xylem_b_int` \times \text{VAF}^{`r vaf_xylem_b_power`}
$$

```{r}

withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_summary_segments_noxylem_traits_log_vaf))

withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_summary_segments_xylem_traits_log_vaf))

withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit3_summary_segments_noxylem_traits_sp_log_vaf))

withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit4_summary_segments_noxylem_traits_sp2_log_vaf))
```



```{r}
d <- read_csv(here::here("data/fd_k_traits.csv")) |>
  filter(species == "Hevea brasiliensis") |>
  dplyr::select(species, sample_number, ks, vaf)

# d |> kable()

d_samp <- d |>
  group_by(sample_number) |>
  summarise(ks = mean(ks), vaf = mean(vaf, na.rm = TRUE))

d_samp

mean_ks <- mean(d$ks)
mean_vaf <- mean(d$vaf, na.rm = TRUE)
mean_ks

s1 <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_summary_segments_noxylem_traits_log_ks))
draws1 <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_draws_segments_noxylem_traits_log_ks)) |>
    janitor::clean_names()

s2 <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_summary_segments_xylem_traits_log_ks))
draws2 <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit2_draws_segments_xylem_traits_log_ks)) |>
    janitor::clean_names()

s1 |>
  filter(str_detect(variable, "^alpha")) |>
  filter(str_detect(variable, "15")) |>
  mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  kable()

draws1 |>
  dplyr::select(starts_with("alpha")) |>
  dplyr::select(ends_with("15"))

s2 |>
  filter(str_detect(variable, "^alpha")) |>
  filter(str_detect(variable, "15")) |>
  mutate(across(where(is.numeric), \(x) round(x, 2))) |>
  kable()

model4 <- withr::with_dir(rprojroot::find_root('_targets.R'),
    targets::tar_read(fit_draws_segments_xylem_0.08)) |>
    janitor::clean_names()

model4 |>
  dplyr::select(starts_with("alpha")) |>
  dplyr::select(ends_with("15"))

```
No xylem or with xylme doesn't matter. I don't need to fix the following:


- a from ks, sp
- a from ks, xylem
- b from ks, no xylem
- b from ks, xylem
- a from vaf, no xylem
- a from vaf, xylem
- b from vaf, no xylem
- b from vaf, xylem

- a from model4, pg = 0.08
- a from model4, pg = 0.07
...


```{r}
m1 <- draws1 |>
  dplyr::select(c(alpha_a_1_15, alpha_a_2_15)) |>
  as.matrix()
m2 <- draws2 |>
  dplyr::select(c(alpha_a_1_15, alpha_a_2_15)) |>
  as.matrix()
```

a from ks, no xylem

```{r}
a_ks_post <- m1 %*% c(1, log(mean_ks))
exp(a_ks_post) |> quantile(c(0.025, 0.5, 0.975))
```

a from ks, xylem

```{r}
a_ks_post <- m2 %*% c(1, log(mean_ks))
exp(a_ks_post) |> quantile(c(0.025, 0.5, 0.975))
```

a from model4, pg = 0.08

```{r}
a_post <- model4 |> pull(alpha_1_15) |> exp()  |> quantile(c(0.025, 0.5, 0.975))
a_post
```
