---
title: "Habits"
author: "Masatoshi Katabuchi"
format:
  revealjs:
    chalkboard: true
    theme: [default]
---

```{r global_options, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  cache = FALSE,
  fig.align = "center",
  fig.show = "hold",
  root.dir = rprojroot::find_root('_targets.R')
)
```


```{r,include=FALSE}
library(tidyverse)
library(tarchetypes)
library(here)
library(kableExtra)
library(targets)
library(loo)
library(smatr)
library(GGally)
library(ggridges)
library(ggrepel)
library(bayesplot)
source(here("R", "stan.R"))
source(here("R", "figs.R"))
anova_yml <- yaml::yaml.load_file(here("yml/anova.yml"))
```

# Fig. 1


**Fig. 1.**
![Flow rate under three levels of driving force employed as pressure or tension, normalized to maximum hydraulic conductivity (Ks) to account for small differences in sample lengths and diameter demonstrating the effect of increasing force on Ks (??).
(A) Relationship between sample mean pressure- and tension-induced flow rates.
Blue solid line indicates a standardized major axis (SMA) regression.
The 95% confidence interval is represented as the shaded area.
(B) Comparisons between pressure- and tension-induced flow rates at three levels for three diffuse-porous, one ring-porous, and one liana species.
The center line in each box indicates the median, and upper and lower box sides indicate the interquartile range.
The whiskers extend to a maximum of 1.5 times the interquartile range.
This is only to simplify the visualization, and statistical results are shown in SX.
HH, *Hopea hongayensis*;
VM, *Vatica mangachapoi*;
HB, *Hevea brasiliensis*;
TG, *Tectona grandis*;
AP, *Acacia pennata*.](`r here::here("figs/sma_ks.png")`)

---


# Fig.

hello
